---
title: Marine mollusc diversity, Galiano Island, BC
author: Antranig Basman, Karin Fletcher,  Charles J. Gibbs, Rick M. Harbo,  Andy Lamb, Andrew Simon
date: April 22nd, 2023
output: html_document
knit: (function(inputFile, encoding) {
  rmarkdown::render(inputFile, encoding = encoding, output_dir = "docs") })
---

```{r, echo=FALSE, message=FALSE, warning=FALSE}
library(tidyverse)
molluscs <- read.csv("tabular_data/Galiano_marine_animal_records_consolidated_2023-04-22.csv")
mollusc.summary <- read.csv("tabular_data/mollusc_summary.csv")
new <- mollusc.summary %>% filter(str_detect(reportingStatus, "new")) 
confirmed <- mollusc.summary %>% filter(reportingStatus == 'confirmed') 
reported <- mollusc.summary %>% filter(reportingStatus == 'reported') 
observed <- mollusc.summary %>% filter(iNatObservationStatus == 'observed') 
unobserved <- mollusc.summary %>% filter(iNatObservationStatus == 'unobserved') 
```

## Galiano Island's marine mollusc diversity

With about 49,000 extant species, Mollusca are the second largest phylum after Arthropoda. The phylum comprises seven extant classes, five of which are locally represented, including bivalves, cephalopods, gastropods, chitons, and tusk shells. Over 780 marine species are been identified in British Columbia, `r nrow(mollusc.summary)` of which have been reported for Galiano Island over the last century and a half, through the combined efforts of `r length(unique(molluscs$recordedBy))` individuals.

```{r, echo=FALSE, message=FALSE, warning=FALSE}
source("scripts/Marine_mollusc_map_script.R")
animal.grid.map
```

## Community science contributions

Since the Biodiversity Galiano project began in 2016, our community has documented `r nrow(observed)` mollusc species in the waters around Galiano Island, confirming `r round((nrow(confirmed)/(nrow(confirmed)+nrow(reported))*100), digits = 0)` percent of the historical records and adding `r nrow(new)` new species to the list. Of the `r nrow(mollusc.summary)` species reported for the island, `r nrow(unobserved)` species have not been seen in the last twenty years.

```{r, echo=FALSE, message=FALSE, warning=FALSE}
source("scripts/Community_science_contributions.R")
reportingStatusMap
reportingStatusFig
```

## Top community contributions to our knowledge of the island mollusc diversity

Browse photos of some of the most commonly documented mollusc species, as well as recent observations and top observers, based on [iNaturalist](http://inaturalist.org/) data. You can help contribute to the growing record of the island mollusc diversity by submitting your observations to the [Biodiversity Galiano iNaturalist project.](http://inaturalist.org/projects/biodiversity-galiano-island) 

<span class="mxcw-nonLeafletPane mxcw-paneName-TopContributions html-widget leaflet"></span>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
